\section{Scope of the Work}
This thesis focuses on the recognition of ISL gestures from RGB video streams using a deep learning pipeline that combines spatial feature extraction, temporal modeling, and attention-based feature refinement. The implementation scope includes data preprocessing, model design, training and inference workflow, and output generation as text and speech. A key requirement within this scope is real-time oriented performance, so architectural choices are made with computational efficiency in mind rather than only maximizing offline benchmark accuracy.

The present work is intentionally bounded to maintain technical clarity and measurable outcomes. The system is developed primarily for isolated gesture recognition and structured evaluation under controlled experimental settings. While the framework can inform future extensions, full continuous sentence-level translation, multilingual sign integration, and edge deployment on low-power embedded hardware are treated as beyond the immediate scope of this thesis and are discussed as forward directions.

Within this scope, certain assumptions are maintained for experimental consistency. The input stream is assumed to provide sufficiently visible upper-body gesture regions, and training data is assumed to represent the target sign classes with reasonable class balance. These assumptions help isolate modeling and inference behavior, while still preserving practical relevance for future real-time deployment refinement.
