\section{Overview of Existing Systems}
Sign language research has evolved from handcrafted-feature pipelines to large-scale deep learning systems. Early recognition systems were typically constrained to small vocabularies, controlled capture settings, or handcrafted descriptors, and therefore struggled to generalize to real communication scenarios. The deep learning transition, anchored by representation learning advances \cite{lecun2015deep}, enabled end-to-end feature discovery from visual data and significantly improved recognition quality.

Modern systems are now commonly organized along two dimensions: task complexity and modality design. In terms of task complexity, the field progressed from isolated sign recognition to continuous sign language recognition (CSLR), and then to full sign language translation (SLT), where sign sequences are mapped to spoken-language sentences with different grammar and word order \cite{camgoz2018neural,camgoz2020transformers}. In terms of modality design, approaches range from RGB-only pipelines to multimodal pipelines that combine RGB with pose, depth, or linguistic priors for better robustness.

Dataset growth has been central to this progress. Large public benchmarks such as MS-ASL and WLASL improved word-level recognition research \cite{joze2019msasl,li2020wlasl}, while RWTH-PHOENIX-Weather and How2Sign supported continuous recognition and translation benchmarking \cite{camgoz2018neural,duarte2021how2sign}. More recent resources, such as Sem-Lex and iSign, further expanded coverage through phonological annotations and large ISL-aligned text-video pairs \cite{kezar2023semlex,joshi2024isign}. This dataset expansion has shifted the field from proof-of-concept models toward scalable and reproducible evaluation.

From a systems perspective, this evolution also reflects a shift from model-centric design to pipeline-centric design. Earlier studies primarily reported top-line accuracy, whereas recent work increasingly reports end-to-end behavior including preprocessing stability, temporal consistency, and inference-time feasibility. As a result, current literature is better viewed as an ecosystem of interconnected components (data, feature extraction, temporal modeling, alignment, and decoding) rather than isolated architecture improvements.

Another notable trend is the growing emphasis on reproducibility and standardized reporting. As datasets become larger and methods become more complex, fair comparison now requires explicit discussion of training splits, preprocessing assumptions, and inference settings. This methodological transparency is increasingly treated as a core requirement for progress in CSLR and ISL-specific research.

At the same time, recent surveys report that high offline performance does not automatically guarantee deployment readiness. Generalization across signers, domains, and devices remains challenging, and real-time constraints continue to expose gaps in latency, robustness, and practical usability \cite{alqurishi2021deep,alabdullah2024advancements}. Therefore, existing systems should be understood as strong but still transitional steps toward reliable real-time assistive communication.
