\section{Existing Methods and Techniques}
Current sign language systems are predominantly built as spatiotemporal learning pipelines. The spatial stage encodes hand shape, orientation, and local appearance cues, often through transfer-learning backbones or 2D/3D convolutional models. Recent work has shown that strong visual backbones such as EfficientNet can improve representation quality when combined with task-specific contextual refinement \cite{yenisari2025deep}. The temporal stage then models sequential dependencies using recurrent architectures (LSTM/GRU), temporal convolutions, or transformer encoders, depending on sequence length, latency constraints, and training scale \cite{camgoz2020transformers,zhou2024transformer}.

A second major direction is sequence alignment and decoding. Since frame-level sign boundaries are hard to annotate at scale, many CSLR systems use CTC-based alignment for weakly aligned supervision. This strategy has become practical for real-time and continuous settings because it supports variable-length input and output sequences. The SignFlow framework, for example, combines pretrained feature extraction, transformer-based temporal learning, and CTC fine-tuning, while introducing device-aware down-sampling and detection-rate analysis for real-time operation in ISL scenarios \cite{geetha2025toward}.

A third direction is multimodal and attention-guided fusion. RGB captures rich appearance information, whereas pose and landmark streams offer geometric motion cues that are often less sensitive to background variation. Attention mechanisms are increasingly used to weight these feature streams adaptively, improving robustness for subtle gestures and speed variation \cite{yenisari2025deep,geetha2025toward}. Gloss-aware and gloss-free translation models further extend this line by integrating local semantic boundary modeling and sentence-level language priors \cite{yin2023gaslt,zhang2024scaling}.

In practical deployments, these method families are rarely used in isolation. Competitive systems often combine transfer-learned visual backbones, lightweight temporal encoders, and alignment-aware decoders in a staged training strategy. This combination allows models to retain discriminative spatial features while still controlling sequence-level drift and runtime overhead, which is especially important for continuous ISL usage in assistive settings.

Finally, dataset-centric methods and pretraining strategies now play a critical role in model quality. Large-scale supervised resources such as iSign and Sem-Lex provide stronger linguistic and phonological coverage \cite{joshi2024isign,kezar2023semlex}, while broad multimodal datasets such as How2Sign improve generalizable sequence modeling \cite{duarte2021how2sign}. Recent large-scale unification efforts suggest that cross-dataset pretraining and shared representations can further improve transferability across sign languages and downstream tasks \cite{li2025unisign}. Overall, the field has shifted from single-model optimization to joint optimization of data curation, architecture design, and alignment-aware training.

Table~\ref{tab:paper-method-comparison} summarizes representative systems reviewed in this chapter using a uniform comparison format.

\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\textwidth}{|p{3.1cm}|p{3.3cm}|X|X|}
\hline
\textbf{Paper / System} & \textbf{Method Used} & \textbf{Advantage} & \textbf{Limitation} \\ \hline
Yenisari and Yavuz (2025) \cite{yenisari2025deep} & EfficientNet + CNN/RNN + attention (isolated SLR) & Strong contextual feature learning and high benchmark accuracy & RGB-centric setup; higher runtime cost for large models \\ \hline
Geetha et al. (2025) SignFlow \cite{geetha2025toward} & CNN + Transformer + CTC with RGB + pose (continuous ISL) & Real-time-oriented design with improved WER and domain-specific pretraining & Requires careful pretraining and runtime tuning \\ \hline
Camgoz et al. (2020) \cite{camgoz2020transformers} & Joint transformer-based end-to-end CSLR/SLT & Unified recognition and translation within one sequence model & Needs large aligned corpora; weaker portability in low-resource settings \\ \hline
Joshi et al. (2024) iSign \cite{joshi2024isign} & Large-scale ISL benchmark with multi-task baselines & Major improvement in ISL data scale and evaluation standardization & Primarily benchmark-focused, not an end-to-end deployment system \\ \hline
Kezar et al. (2023) Sem-Lex \cite{kezar2023semlex} & SL-GCN with phonological auxiliary targets & Better few-shot recognition and improved linguistic interpretability & Isolated ASL focus; limited direct coverage for continuous translation \\ \hline
\end{tabularx}
\caption{Representative SLR systems: methods and limits}
\label{tab:paper-method-comparison}
\end{table}

The comparison indicates a consistent pattern: methods that balance representation quality, temporal alignment, and computational efficiency perform better for practical ISL usage than methods that optimize only one dimension. This reinforces the need for integrated architecture design and motivates the methodology proposed in this thesis.
